---
title: "Vowel duration and geminate stops in Icelandic"
author: "Stefano Coretta"
date: "19/04/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
library(tidyverse)
theme_set(theme_light())
library(lmerTest)
library(effects)
```

# Read data

```{r data}
words <- read_csv("task/task-src/words.csv")

stops <- read_csv(
    "results/results.csv",
    na = "--undefined--"
    ) %>%
    right_join(y = words) %>%
    mutate(
        syllables = factor(syllables, c("mono","di")),
        vowel = factor(vowel, c("i", "ɪ", "ʏ", "e", "œ", "a", "o", "ei", "ai", "ou")),
        syl_structure = factor(syl_structure, c("cvcc", "vcc", "cvvcc", "vvcc", "cvccv", "cvvccv"))
        ) %>%
    filter(syl_structure %in% c("cvcc", "cvccv")) %>%
    mutate_if(is.character, as.factor) %>%
    filter(manner == "stop", contx == "cc") %>%
    droplevels()
```

# Exploratory analysis

```{r}
stops %>%
    ggplot(aes(word)) + geom_bar()
```

```{r}
stops %>%
    ggplot(aes(syllables, fill = aspiration)) + geom_bar()
```

```{r}
stops %>%
    ggplot(aes(syl_structure, fill = aspiration)) + geom_bar()
```

```{r}
stops %>%
    ggplot(aes(word.dur, colour = aspiration)) +
    geom_density() +
    facet_grid(. ~ syl_structure)
```

```{r word-duration}
stops %>%
    ggplot(aes(aspiration, word.dur)) +
    geom_boxplot() +
    facet_grid(. ~ syl_structure)
```

```{r}
stops %>%
    ggplot(aes(aspiration, word.dur)) +
    geom_boxplot() +
    facet_grid(syl_structure ~ vowel)
```

```{r}
stops %>%
    ggplot(aes(aspiration, vor)) +
    geom_boxplot() +
    facet_grid(. ~ syl_structure)
```

## Vowel duration

```{r}
stops %>%
    ggplot(aes(aspiration, vowel.dur)) +
    geom_boxplot() +
    facet_grid(. ~ syl_structure)
```

```{r}
stops %>%
    ggplot(aes(syl_structure, vowel.dur)) +
    geom_violin() +
    geom_boxplot(width = 0.2)
```


```{r}
stops %>%
    ggplot(aes(aspiration, vowel.dur)) +
    geom_boxplot() +
    facet_grid(syl_structure ~ vowel)
```

```{r}
stops %>%
    ggplot(aes(aspiration, vowel.dur)) +
    geom_boxplot() +
    facet_wrap(~ vowel)
```

```{r}
stops %>%
    ggplot(aes(closure.dur, vowel.dur, colour = aspiration)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm")
```


## Voicing duration

```{r}
stops %>%
    ggplot(aes(aspiration, voicing.dur)) +
    geom_boxplot() +
    facet_wrap(~ vowel)
```

# LMER

## Vowel duration

```{r vowel-lmer}
vowel.lmer <- lmer(
    vowel.dur ~
        aspiration *
        syl_structure +
        vowel +

        c1phonation +
        vor +
        (1|word) +
        (1+aspiration|speaker),
    data = stops,
    REML = FALSE
)

summary(vowel.lmer)

vowel.lmer.null <- lmer(
    vowel.dur ~
        vowel +
        syl_structure +
        c1phonation +
        vor +
        (1|word) +
        (1|speaker),
    data = stops,
    REML = FALSE
)

summary(vowel.lmer.null)

vowel.lmer.null.2 <- lmer(
    vowel.dur ~
        aspiration +
        vowel +
        syl_structure +
        c1phonation +
        vor +
        (1|word) +
        (1|speaker),
    data = stops,
    REML = FALSE
)
```

```{r vowel-lmer-anova}
anova(vowel.lmer.null, vowel.lmer)

anova(vowel.lmer.null.2, vowel.lmer)
```

The best model is `vowel.lmer`. There is an iteraction between aspiration and syllable structure: the effect of aspiration is stronger in CVCC words then in CVCCV words.

```{r vowel-lmer-plot}
plot(allEffects(vowel.lmer))
```

```{r}
plot(fitted(vowel.lmer), resid(vowel.lmer))
```

## Word duration

```{r word-lmer}
word.lmer <- lmer(
    word.dur ~
        aspiration *
        syl_structure +
        vowel +
        c1phonation +
        vor +
        (1|word) +
        (1|speaker),
    data = stops
)

summary(word.lmer)
```

```{r word-lmer-plot}
plot(allEffects(word.lmer))
```

CVCC words are shorter if CC is pre-aspirated, while this does not happen with CVCCV words.
However, CVCCV words show a bimodal distribution I don't know how to explain.